### 2019-03-30 Thesis/games.tex
D.

\documentclass{article}

@@

\section{A Fully Abstract Game Semantics for Idealized Algol}

@@

\subsection{Idealized Algol}

@@

\begin{lstlisting}[language=Java, morekeywords={nat,com}]
public interface Var
{
  nat read();
  com write(nat value);
}
\end{lstlisting}

@@

\subsection{Games and Strategies}

@@

\subsection{Connectives on Games}
\label{SecConnectives}

@@

\begin{definition}
  Given games $A$,$B$, define a game $A\times B$ by
  \begin{itemize}
    \item $M_{A\times B} = M_A + M_B$,
    \item $\lambda_{A\times B} = [\lambda_A,\lambda_B]$,
    \item $* \ts_{A\times B} a$ if and only if $*\ts_A a$ or $*\ts_B a$ and $a \ts_{A\times B} b$ if and only if $a \ts_A b$ or $a\ts_B b$, and
    \item $P_{A\times B} = \{s\in L_{A\times B} \suchthat \text{$s\vert_A\in P_A$ and $s\vert_B=\epsilon$ or $s\vert_A=\epsilon$ and $s\vert_B\in P_B$}\}$.
  \end{itemize}
  We extend this to arbitrary products $\prod_i A_i$ in the obvious way.
  In particular, the product $1$ of the empty collection is the same as the empty game $I$ defined in Example \ref{ExEmptyGame}.
  \label{DefProduct}
\end{definition}

@@

\subsection{Composition of strategies}

@@

\begin{definition}[{\cite[\sec 3.1]{Harmer2006InnocentGS}}]
  Given $\s\in \Int(A,B,C)$, we define the \emph{$P$-view} $\pv{\s}$ of $\s$ inductively as follows.
  \begin{IEEEeqnarray*}{RCL?s}
    \pv{\epsilon} & = & \epsilon & \\
    \pv{\s a} & = & \pv{\s}a & if $a$ is a move in $B$, an $O$-move in $A$ or a $P$-move in $C$ \\
    \pv{\s c} & = & c & if $c$ is an initial move of $C$ \\
    \pv{\s b\t a} & = & \pv{\s}ba & if $a$ is a $P$-move of $A$ or an $O$-move of $C$ and is justified by $b$ \\
  \end{IEEEeqnarray*}
\end{definition}

@@

  Having established that $\s\vert_{A,C}$ is alternating, we show that it is well-bracketed.
  Suppose that a question move $q$ in $\s\vert_{A,C}$ justifies some answer move $a$.
  $q$ and $a$ must occur in the same component, since the only case in which a move from one of $A$ and $C$ can justify a move in the other is when both moves are initial, and hence questions.
  Suppose first that $q$ and $a$ both occur in the game $C$.  
  Suppose that some other question move $q'$ occurs between $q$ and $a$ in $\s\vert_{A,C}$.  
  If $q'$ occurs in $C$, then it must be answered by some $a'$ occurring between $q'$ and $a$, since $\s\vert_C$ is a well-bracketed sequence.  
  Otherwise, suppose that $q'$ occurs in $A$.

@@

\subsection{Associativity of composition}

In this section, we will prove that composition is associative; i.e., that if $\sigma\from A \implies B$, $\tau\from B\implies C$ and $\upsilon \from C \implies D$ are strategies, then $(\sigma;\tau);\upsilon=\sigma;(\tau;\upsilon)$.  
To do this, if $A,B,C,D$ are arenas, we define the set $\Int(A,B,C,D)$ to be the set of all sequences $\u$ of moves such that $\u\vert_{A,B}\in L_{A\implies B}$, $\u\vert_{B,C}\in L_{B\implies C}$ and $\u\vert_{C,D}\in L_{C \implies D}$.  
Given such a sequence $\u$, we define $\u\vert_{A,D}$ as before; i.e., we take all moves from $\u$ occurring in $A$ and $D$, together with justification pointers within these games, and if an initial move in $A$ is justified by an initial move in $B$, which is justified by an initial move in $C$, which is justified by an initial move in $D$, then we add a justification pointer from that move in $A$ to that move in $D$.

@@

\subsection{Copycat strategies}
\label{SecCopycat}

@@

It is also easy to see from Proposition \ref{PropCopycat} that the identity function $\id\from M_A \to M_A$ is a structural isomorphism from $A$ to itself, and that the resulting copycat strategy $\cc_{\id}$ is an identity for composition.  
Combining this with our result for associativity in the previous section, we get that
\begin{theorem}
  The collection of games forms a category $\G$, where the morphisms $A \to B$ are strategies for $A\implies B$, composition is as above and the identity morphisms are the copycat strategies induced from the identity functions on moves.
\end{theorem}

@@

\subsection{$\G$ as a Symmetric Monoidal Category}

@@

The above proof gives us the following result analogous to Corollary \ref{CorSwitchingCondition}.

@@

  Now suppose that $sab,sac\in\sigma\tensor\tau$.  
  Then $sa$ is an $O$-position in $P_{(A\tensor C)\implies (B\tensor D)}$, and is therefore in configuration $PPPO$ or $PPOP$.  
  By inspecting the table above, we see that if $sa$ is in configuration $PPPO$, then $b$ and $c$ must both occur either in $C$ or in $D$, and that if $sa$ is in configuration $PPOP$, then $b$ and $c$ must both occur either in $A$ or in $B$.  
  In either case, we must have $b=c$, by determinism of $\tau$ (in the first case) or of $\sigma$ (in the second case).
\end{proof}

@@

\subsection{$\G$ as a Symmetric Monoidal Closed Category}

@@

  We examine the sign configuration of a play in $(B\implies C)\implies (A\implies D)$, using Lemma \ref{LemCompositionLemma}.  
  Since we must avoid the configuration $OP$ in either $B\implies C$, $A\implies D$ or in $(B\implies C)\implies (A\implies D)$, we arrive at the following list of possibilities.
  \scriptsize
  \[
\arraycolsep=2.7pt\def\arraystretch{1.5}
    \begin{array}{cccc|cc|c}
      \lambda_B^{OP}(s\vert_B) & \lambda_C^{OP}(s\vert_C) & \lambda_A^{OP}(s\vert_A) & \lambda_D^{OP}(s\vert_D) & \lambda_{B\implies C}^{OP}(s\vert_{B,C}) & \lambda_{A\implies D}^{OP}(s\vert_{A,D}) & \lambda_{(B\implies C)\implies (A\implies D)}^{OP}(s) \\
      \hline
      P & O & P & O & O & O & P \\
      P & P & P & O & P & O & O \\
      O & O & P & O & P & O & O \\
      P & P & P & P & P & P & P \\
      O & O & O & O & P & P & P \\
      P & P & O & O & P & P & P \\
      O & O & P & P & P & P & P
    \end{array}
    \]
  \normalsize
  If $s\in\sigma\implies\tau$, then we can immediately discount the last two of these possibilities, since one includes the illegal configuration $OP$ in $A\implies B$, and the other includes the illegal configuration $OP$ in $B\implies D$.

@@

\subsection{Products in $\G$}

@@

\subsection{Sequoidal categories}

We have now given the category-theoretic properties of all the connectives from Section \ref{SecConnectives}, with the exception of the sequoid $\sequoid$ and the exponential $\oc$.

@@

\begin{definition}
  A \emph{sequoidal category} $\C$ is given by
  \begin{itemize}
    \item a monoidal category $(\C,\tensor, I)$ (with coherences $\assoc,\lunit,\runit$);
    \item a (strong) right action of $\C$ on a category $\C_s$; i.e., a functor $\blank\sequoid\blank\from\C_s\times\C\to\C_s$ together with natural isomorphisms
      \begin{mathpar}
        \passoc_{a,x,y}\from (a \sequoid x) \sequoid y \toisom a \sequoid (x \tensor y)
        \and
        \run_a \from a \toisom a \sequoid I
      \end{mathpar}
      that make the diagrams
      \begin{mathpar}
        \begin{tikzcd}[column sep=36pt]
          ((a \sequoid x) \sequoid y) \sequoid z \arrow[r, "{\passoc_{a,x,y}\sequoid z}"] \arrow[d, "{\passoc_{a\sequoid x,y,z}}"' description]
            & (a \sequoid (x \tensor y)) \sequoid z \arrow[r, "{\passoc_{a,x\tensor y,z}}"]
              & a \sequoid ((x\tensor y) \tensor z)  \arrow[dl, "{a \sequoid \assoc_{x,y,z}}"] \\
          (a \sequoid x) \sequoid (y \tensor z) \arrow[r, "{\passoc_{a,x,(y\tensor z)}}"]
            & a \sequoid (x \tensor (y \tensor z))
              &
        \end{tikzcd}
        \and
        \begin{tikzcd}
          a \sequoid x \arrow[r, "{a \sequoid \lunit_x}"] \arrow[d, "\run_a \sequoid x"']
            & a \sequoid (I \tensor x) \\
          (a \sequoid I) \sequoid x \arrow[ur, "{\passoc_{a,I,x}}"']
            &
        \end{tikzcd}
        \and
        \begin{tikzcd}
          a \sequoid x \arrow[r, "{a \sequoid \runit_x}"] \arrow[d, "\run_{a \sequoid x}"']
            & a \sequoid (x \tensor I) \\
          (a \sequoid x) \sequoid I \arrow[ur, "{\passoc_{a,x,I}}"']
            &
        \end{tikzcd}
      \end{mathpar}
      commute; and
    \item a lax morphism of actions from $\blank\sequoid\blank$ to the right tensor multiplication action $\blank\tensor\blank$ of $\C$ on itself; i.e., a functor $J\from \C_s \to \C$ and a natural transformation $\wk_{a,x} \from Ja \tensor x \to J(a \sequoid x)$ that makes the following diagrams commute.
      \begin{mathpar}
        \begin{tikzcd}[column sep=30pt]
          (Ja \tensor x) \tensor y \arrow[r, "{\wk_{a,x}\tensor y}"] \arrow[d, "{\assoc_{Ja,x,y}}"'] & J(a \sequoid x) \tensor y \arrow[r, "{\wk_{a\sequoid x,y}}"]
              & J((a \sequoid x) \sequoid y) \arrow[dl, "{J\passoc_{a,x,y}}"] \\
          Ja \tensor (x \tensor y) \arrow[r, "{\wk_{a,x\tensor y}}"]
            & J(a \sequoid (x \tensor y))
              &
        \end{tikzcd}
        \and
        \begin{tikzcd}
          Ja \arrow[r, "J\run_a"] \arrow[d, "\runit_{Ja}"']
            & J(a \sequoid I) \\
          Ja \tensor I \arrow[ur, "{\wk_{a,I}}"']
            &
        \end{tikzcd}
      \end{mathpar}
  \end{itemize}
  \label{DefSequoidalCategory}
\end{definition}
\begin{remark}
  The definitions of \emph{lax action} can be found at the start of Section \ref{SecParametricMonads}, whie that of an \emph{oplax morphism of actions} is found at Definition \ref{DefOplaxMorphismOfActions}.  
  The definitions we have used are similar: a \emph{strong action} is a lax action in which the coherences (called $m$ and $e$ in Section \ref{SecParametricMonads} and $\passoc$ and $\run$ here) are isomorphims.  
  A \emph{lax morphism of actions} is defined in the same way as an oplax morphism, except that the coherence (called $\mu$ in Definition \ref{DefOplaxMorphismOfActions} and $\wk$ here) goes in the opposite direction.
\end{remark}

@@

  Let us start with the diagrams.  
  By Proposition \ref{PropCopycat}, commutativity of these diagrams follows from commutativity of the diagrams formed from the corresponding subset inclusion functions in $\Set$.  
  For example, to show that the first diagram commutes in $\G$, we must show that the following diagram commutes in $\Set$.
  \scriptsize
  \[
    \begin{tikzcd}[column sep=19pt, row sep=30pt]
      ((M_A + M_X) + M_Y) + M_Z \arrow[r, "{[\assoc_{M_A,M_X,M_Y},\id_{M_Z}]}" yshift=3pt] \arrow[d, "{\assoc_{M_A+M_X,M_Y,Z}}" description]
        & (M_A + (M_X + M_Y)) + M_Z \arrow[r, "{\assoc_{M_A,M_X+M_Y,M_Z}}" yshift=3pt]
          & M_A + ((M_X + M_Y) + M_Z) \arrow[dl, "{[\id_{M_A},\assoc_{M_X,M_Y,M_Z}]}"] \\
      (M_A + M_X) + (M_Y + M_Z) \arrow[r, "{\assoc_{M_A,M_X,M_Y+M_Z}}" yshift=3pt]
        & M_A + (M_X + (M_Y + M_Z))
          &
    \end{tikzcd}
    \]
  \normalsize
  This diagram is, of course, none other than the pentagram diagram for the coproduct $+$ in $\Set$.  
  Similarly, the second and third diagrams in Definition \ref{DefSequoidalCategory} reduce in this case to the triangle diagrams for the coproduct $+$ in $\Set$.

  For the fourth diagram in Definition \ref{DefSequoidalCategory}, since $\wk$ is a subset inclusion strategy induced from an identity map, Proposition \ref{PropCopycat} tells us that both arms of the diagram are the strategy induced by the subset inclusion $\assoc_{M_A,M_X,M_Y}\from (M_A + M_X) + M_Y \to M_A + (M_X + M_Y)$ from $(A \tensor X) \tensor Y$ to $A \sequoid (X \tensor Y)$.
  Similarly, both arms of the last diagram in Definition \ref{DefSequoidalCategory} are the strategies induced by the subset inclusion $\runit_{M_A} \from M_A \to M_A + \emptyset$ from $A$ to $A \sequoid I$.

@@

\subsection{Tree Embeddings and Zigzag Strategies}

@@

We can generalize the subset inclusions of the previous section to \emph{tree embeddings}.  
Tree embeddings are similar to subset inclusions, but generated by a function between plays, rather than between moves.  
A consequence of this is that while tree embeddings do give rise to strategies, these strategies are not in general innocent.

@@

\subsection{Sequoidally decomposable categories}

@@

\begin{proposition}
  $\G$ is a strongly distributive sequoidal category.
\end{proposition}
\begin{proof}
  By Proposition \ref{PropProductOfTreeEmbeddings}, the morphism $\langle (A_1\sequoid(\cdots\sequoid(A_n \sequoid\pr_i)\cdots))\rangle$ is given by the tree embedding $\phi\from P_{\prod_i (A_1 \sequoid(\cdots\sequoid(A_n\sequoid B_i)\cdots))} \to P_{A_1\sequoid(\cdots\sequoid(A_n\sequoid \prod_iB_i))}$ defined as follows.
  \[
    \phi(s) = \begin{cases}
      \epsilon & \text{if $s=\epsilon$} \\
      [\inj_{A_1,\cdots,A_n},\inj_{B_j}]_*(s) & \text{if $s$ begins with a move in the $j$-th component}
    \end{cases}
    \]
  Note that $\phi$ is not in general injective, since if $s$ occurs entirely inside one of the copies of the $A_i$, then $\phi(s)=\phi(s')$ for any identical sequence $s'$ occurring inside one of the other copies of the $A_i$.

@@

\subsection{A Formula for the Exponential}

@@

\begin{theorem}[\cite{MelliesCofCommCom}]
  Let $\C$ be a monoidal category in which the monoidal unit is a terminal object.
  Suppose that $\C$ has symmetrized tensor powers that commute with the tensor product.

@@

  Then the canonical morphisms $B \tensor A^{\sequoid (n+1)}\to B\tensor A^{\sequoid n}$ are given by
  \[
    B \tensor (A \sequoid \blank)^n()\,,
    \]
  where $() \from A \to I$ is the unique morphism into the terminal object.
\end{proposition}
\begin{proof}
  First, we show by induction on $n$ that the following diagram commutes.
  \[
    \begin{tikzcd}[column sep=40pt]
      A^{\tensor(n+1)} \arrow[r, "A^{\tensor n}\tensor()"] \arrow[d, "A\tensor\wk^{n+1}"']
        & A^{\tensor n} \arrow[d, "\wk^n"] \\
      A^{\sequoid(n+1)} \arrow[r, "(A\sequoid\blank)^n()"]
        & A^{\sequoid n}
    \end{tikzcd}
    \]
  This is clearly true for $n=0$; in the general case, we have the following commutative diagram --
  \[
    \begin{tikzcd}[column sep=42pt]
      A^{\tensor(n+2)} \arrow[r] \arrow[dd, "\wk^{n+2}"']
        & A \tensor A^{\tensor (n+1)} \arrow[r, "A \tensor (A^{\tensor n}\tensor ())"] \arrow[d, "A \tensor \wk^{n+1}"']
          & A \tensor A^{\tensor n} \arrow[r] \arrow[d, "A\tensor \wk^n"]
            & A^{\tensor(n+1)} \arrow[dd, "\wk^{n+1}"] \\
      %
        & A \tensor A^{\sequoid (n+1)} \arrow[r, "A \tensor (A\sequoid\blank)^n()"] \arrow[dl, "{\wk_{A,A^{\sequoid (n+1)}}}" description]
          & A \tensor A^{\sequoid n} \arrow[dr, "{\wk_{A,A^{\sequoid n}}}" description]
            & \\
      A^{\sequoid(n+1)} \arrow[rrr, "(A \sequoid \blank)^{n+1}()"]
        &
          &
            & A^{\sequoid (n+1)}
    \end{tikzcd}\,\text{--}
    \]
  where the middle square is the inductive hypothesis (tensored by $A$), the outer trapezia are the definitions of $\wk^{n+2}$ and $\wk^{n+1}$, and the bottom trapezium commutes because $\wk$ is a natural transformation.

@@

\subsection{The Exponential as a Final Coalgebra}

@@

\subsection{Denotational Semantics of Idealized Algol}

@@

\subsection{The Strategy $\cell$}

@@

\subsection{Small-Step Operational Semantics of Idealized Algol}

@@

\subsection{Soundness}

>>>>

\documentclass[11pt]{report}

@@

\chapter{A Fully Abstract Game Semantics for Idealized Algol}

@@

\section{Idealized Algol}

@@

\begin{minipage}{\linewidth}
\begin{lstlisting}[language=Java, morekeywords={nat,com}]
public interface Var
{
  nat read();
  com write(nat value);
}
\end{lstlisting}
\end{minipage}

@@

\section{Games and Strategies}

@@

\section{Connectives on Games}
\label{SecConnectives}

@@

\begin{definition}
  Given games $A$,$B$, define a game $A\times B$ by
  \begin{itemize}
    \item $M_{A\times B} = M_A + M_B$,
    \item $\lambda_{A\times B} = [\lambda_A,\lambda_B]$,
    \item $* \ts_{A\times B} a$ if and only if $*\ts_A a$ or $*\ts_B a$ and $a \ts_{A\times B} b$ if and only if $a \ts_A b$ or $a\ts_B b$, and
    \item $P_{A\times B} = \{s\in L_{A\times B} \suchthat \text{$s\vert_A\in P_A$ \& $s\vert_B=\epsilon$ or $s\vert_A=\epsilon$ \& $s\vert_B\in P_B$}\}$.
  \end{itemize}
  We extend this to arbitrary products $\prod_i A_i$ in the obvious way.
  In particular, the product $1$ of the empty collection is the same as the empty game $I$ defined in Example \ref{ExEmptyGame}.
  \label{DefProduct}
\end{definition}

@@

\section{Composition of strategies}

@@

\begin{definition}[{\cite[\sec 3.1]{Harmer2006InnocentGS}}]
  Given $\s\in \Int(A,B,C)$, we define the \emph{$P$-view} $\pv{\s}$ of $\s$ inductively as follows.
  \begin{IEEEeqnarray*}{RCL?s}
    \pv{\epsilon} & = & \epsilon & \\
    \pv{\s a} & = & \pv{\s}a & if $a$ is a move in $B$, an $O$-move in $A$ or a $P$-move in $C$ \\
    \pv{\s c} & = & c & if $c$ is an initial move of $C$ \\
    \pv{\s b\t a} & = & \pv{\s}ba & \parbox[t][][t]{200pt}{if $a$ is a $P$-move of $A$ or an $O$-move of $C$ and is justified by $b$} \\
  \end{IEEEeqnarray*}
\end{definition}

@@

  Having established that $\s\vert_{A,C}$ is alternating, we now show that it is well-bracketed.
  Suppose that a question move $q$ in $\s\vert_{A,C}$ justifies some answer move $a$.
  $q$ and $a$ must occur in the same component, since the only case in which a move from one of $A$ and $C$ can justify a move in the other is when both moves are initial, and hence questions.
  Suppose first that $q$ and $a$ both occur in the game $C$.  
  Suppose that some other question move $q'$ occurs between $q$ and $a$ in $\s\vert_{A,C}$.  
  If $q'$ occurs in $C$, then it must be answered by some $a'$ occurring between $q'$ and $a$, since $\s\vert_C$ is a well-bracketed sequence.  
  Otherwise, suppose that $q'$ occurs in $A$.

@@

\section{Associativity of composition}

In this section, we shall prove that composition of strategies is associative; i.e., that if $\sigma\from A \implies B$, $\tau\from B\implies C$ and $\upsilon \from C \implies D$ are strategies, then $(\sigma;\tau);\upsilon=\sigma;(\tau;\upsilon)$.  
To do this, if $A,B,C,D$ are arenas, we define the set $\Int(A,B,C,D)$ to be the set of all sequences $\u$ of moves such that $\u\vert_{A,B}\in L_{A\implies B}$, $\u\vert_{B,C}\in L_{B\implies C}$ and $\u\vert_{C,D}\in L_{C \implies D}$.  
Given such a sequence $\u$, we define $\u\vert_{A,D}$ as before; i.e., we take all moves from $\u$ occurring in $A$ and $D$, together with justification pointers within these games, and if an initial move in $A$ is justified by an initial move in $B$, which is justified by an initial move in $C$, which is justified by an initial move in $D$, then we add a justification pointer from that move in $A$ to that move in $D$.

@@

\section{Copycat strategies}
\label{SecCopycat}

@@

It is also easy to see from Proposition \ref{PropCopycat} that the identity function $\id\from M_A \to M_A$ is a structural isomorphism from $A$ to itself, and that the resulting copycat strategy $\cc_{\id}$ is an identity for composition.  
Combining this with our result for associativity in the previous chapter, we get that
\begin{theorem}
  The collection of games forms a category $\G$, where the morphisms $A \to B$ are strategies for $A\implies B$, composition is as above and the identity morphisms are the copycat strategies induced from the identity functions on moves.
\end{theorem}

@@

\section{$\G$ as a Symmetric Monoidal Category}

@@

The above proof gives us the following result, which is analogous to Corollary \ref{CorSwitchingCondition}.

@@

  Now suppose that $sab,sac\in\sigma\tensor\tau$.  
  Then $sa$ is an $O$-position in $(A\tensor C)\implies (B\tensor D)$, and is therefore in configuration $PPPO$ or $PPOP$.  
  By inspecting the table above, we see that if $sa$ is in configuration $PPPO$, then $b$ and $c$ must both occur either in $C$ or in $D$, and that if $sa$ is in configuration $PPOP$, then $b$ and $c$ must both occur either in $A$ or in $B$.  
  In either case, we must have $b=c$, by determinism of $\tau$ (in the first case) or of $\sigma$ (in the second case).
\end{proof}

@@

\section{$\G$ as a Symmetric Monoidal Closed Category}

@@

  We examine the sign configuration of a play in $(B\implies C)\implies (A\implies D)$, using Lemma \ref{LemCompositionLemma}.  
  Since we must avoid the configuration $OP$ in either $B\implies C$, $A\implies D$ or in $(B\implies C)\implies (A\implies D)$, we arrive at the following list of possibilities.
  \scriptsize
  \[
\arraycolsep=2.6pt\def\arraystretch{1.5}
    \begin{array}{cccc|cc|c}
      \lambda_B^{OP}(s\vert_B) & \lambda_C^{OP}(s\vert_C) & \lambda_A^{OP}(s\vert_A) & \lambda_D^{OP}(s\vert_D) & \lambda_{B\implies C}^{OP}(s\vert_{B,C}) & \lambda_{A\implies D}^{OP}(s\vert_{A,D}) & \lambda_{(B\implies C)\implies (A\implies D)}^{OP}(s) \\
      \hline
      P & O & P & O & O & O & P \\
      P & P & P & O & P & O & O \\
      O & O & P & O & P & O & O \\
      P & P & P & P & P & P & P \\
      O & O & O & O & P & P & P \\
      P & P & O & O & P & P & P \\
      O & O & P & P & P & P & P
    \end{array}
    \]
  \normalsize
  If $s\in\sigma\implies\tau$, then we can immediately discount the last two of these possibilities, since one includes the illegal configuration $OP$ in $A\implies B$, and the other includes the illegal configuration $OP$ in $B\implies D$.

@@

\section{Products in $\G$}

@@

\section{Sequoidal categories}

We have now given the category-theoretic properties of all the connectives from chapter \ref{SecConnectives}, with the exception of the sequoid $\sequoid$ and the exponential $\oc$.

@@

\begin{definition}[\cite{laird02}]
  A \emph{sequoidal category} $\C$ is given by
  \begin{itemize}
    \item a monoidal category $(\C,\tensor, I)$ (with coherences $\assoc,\lunit,\runit$);
    \item a (strong) right action of $\C$ on a category $\C_s$; i.e., a functor $\blank\sequoid\blank\from\C_s\times\C\to\C_s$ together with natural isomorphisms
      \begin{mathpar}
        \passoc_{a,x,y}\from (a \sequoid x) \sequoid y \toisom a \sequoid (x \tensor y)
        \and
        \run_a \from a \toisom a \sequoid I
      \end{mathpar}
      that make the diagrams
      \begin{mathpar}
        \begin{tikzcd}[column sep=31pt]
          ((a \sequoid x) \sequoid y) \sequoid z \arrow[r, "{\passoc_{a,x,y}\sequoid z}" yshift=1pt] \arrow[d, "{\passoc_{a\sequoid x,y,z}}"' description]
            & (a \sequoid (x \tensor y)) \sequoid z \arrow[r, "{\passoc_{a,x\tensor y,z}}" yshift=1pt]
              & a \sequoid ((x\tensor y) \tensor z)  \arrow[dl, "{a \sequoid \assoc_{x,y,z}}"] \\
          (a \sequoid x) \sequoid (y \tensor z) \arrow[r, "{\passoc_{a,x,(y\tensor z)}}" yshift=1pt]
            & a \sequoid (x \tensor (y \tensor z))
              &
        \end{tikzcd}
        \and
        \begin{tikzcd}
          a \sequoid x \arrow[r, "{a \sequoid \lunit_x}"] \arrow[d, "\run_a \sequoid x"']
            & a \sequoid (I \tensor x) \\
          (a \sequoid I) \sequoid x \arrow[ur, "{\passoc_{a,I,x}}"']
            &
        \end{tikzcd}
        \and
        \begin{tikzcd}
          a \sequoid x \arrow[r, "{a \sequoid \runit_x}"] \arrow[d, "\run_{a \sequoid x}"']
            & a \sequoid (x \tensor I) \\
          (a \sequoid x) \sequoid I \arrow[ur, "{\passoc_{a,x,I}}"']
            &
        \end{tikzcd}
      \end{mathpar}
      commute; and
    \item a lax morphism of actions from $\blank\sequoid\blank$ to the right tensor multiplication action $\blank\tensor\blank$ of $\C$ on itself; i.e., a functor $J\from \C_s \to \C$ and a natural transformation $\wk_{a,x} \from Ja \tensor x \to J(a \sequoid x)$ that makes the following diagrams commute.
      \begin{mathpar}
        \begin{tikzcd}[column sep=30pt]
          (Ja \tensor x) \tensor y \arrow[r, "{\wk_{a,x}\tensor y}"] \arrow[d, "{\assoc_{Ja,x,y}}"'] & J(a \sequoid x) \tensor y \arrow[r, "{\wk_{a\sequoid x,y}}"]
              & J((a \sequoid x) \sequoid y) \arrow[dl, "{J\passoc_{a,x,y}}"] \\
          Ja \tensor (x \tensor y) \arrow[r, "{\wk_{a,x\tensor y}}"]
            & J(a \sequoid (x \tensor y))
              &
        \end{tikzcd}
        \and
        \begin{tikzcd}
          Ja \arrow[r, "J\run_a"] \arrow[d, "\runit_{Ja}"']
            & J(a \sequoid I) \\
          Ja \tensor I \arrow[ur, "{\wk_{a,I}}"']
            &
        \end{tikzcd}
      \end{mathpar}
  \end{itemize}
  \label{DefSequoidalCategory}
\end{definition}
\begin{remark}
  The definitions of \emph{lax action} can be found at the start of chapter \ref{SecParametricMonads}, whie that of an \emph{oplax morphism of actions} is found at Definition \ref{DefOplaxMorphismOfActions}.  
  The definitions we have used are similar: a \emph{strong action} is a lax action in which the coherences (called $m$ and $e$ in chapter \ref{SecParametricMonads} and $\passoc$ and $\run$ here) are isomorphims.  
  A \emph{lax morphism of actions} is defined in the same way as an oplax morphism, except that the coherence (called $\mu$ in Definition \ref{DefOplaxMorphismOfActions} and $\wk$ here) goes in the opposite direction.
\end{remark}

@@

  Let us start with the diagrams.  
  By Proposition \ref{PropCopycat}, commutativity of these diagrams follows from commutativity of the diagrams formed from the corresponding subset inclusion functions in $\Set$.  
  For example, to show that the first diagram commutes in $\G$, we must show that the following diagram commutes in $\Set$.
  \scriptsize
  \[
    \begin{tikzcd}[column sep=16pt, row sep=30pt]
      ((M_A + M_X) + M_Y) + M_Z \arrow[r, "{[\assoc_{M_A,M_X,M_Y},\id_{M_Z}]}" yshift=3pt] \arrow[d, "{\assoc_{M_A+M_X,M_Y,Z}}" description]
        & (M_A + (M_X + M_Y)) + M_Z \arrow[r, "{\assoc_{M_A,M_X+M_Y,M_Z}}" yshift=3pt]
          & M_A + ((M_X + M_Y) + M_Z) \arrow[dl, "{[\id_{M_A},\assoc_{M_X,M_Y,M_Z}]}"] \\
      (M_A + M_X) + (M_Y + M_Z) \arrow[r, "{\assoc_{M_A,M_X,M_Y+M_Z}}" yshift=3pt]
        & M_A + (M_X + (M_Y + M_Z))
          &
    \end{tikzcd}
    \]
  \normalsize
  This diagram is, of course, none other than the pentagram diagram for the coproduct $+$ in $\Set$.  
  Similarly, the second and third diagrams in Definition \ref{DefSequoidalCategory} reduce in this case to the triangle diagrams for the coproduct $+$ in $\Set$.

  For the fourth diagram in Definition \ref{DefSequoidalCategory}, since $\wk$ is a subset inclusion strategy induced from an identity map, Proposition \ref{PropCopycat} tells us that both arms of the diagram are the strategy induced by the subset inclusion 
  \[
    \assoc_{M_A,M_X,M_Y}\from (M_A + M_X) + M_Y \to M_A + (M_X + M_Y)
    \]
  from $(A \tensor X) \tensor Y$ to $A \sequoid (X \tensor Y)$.
  Similarly, both arms of the last diagram in Definition \ref{DefSequoidalCategory} are the strategies induced by the subset inclusion $\runit_{M_A} \from M_A \to M_A + \emptyset$ from $A$ to $A \sequoid I$.

@@

\section{Tree Embeddings and Zigzag Strategies}

@@

We can generalize the subset inclusions of the previous chapter to \emph{tree embeddings}.  
Tree embeddings are similar to subset inclusions, but generated by a function between plays, rather than between moves.  
A consequence of this is that while tree embeddings do give rise to strategies, these strategies are not in general innocent.

@@

\section{Sequoidally decomposable categories}

@@

\begin{proposition}
  $\G$ is a strongly distributive sequoidal category.
\end{proposition}
\begin{proof}
  By Proposition \ref{PropProductOfTreeEmbeddings}, the morphism $\langle (A_1\sequoid(\cdots\sequoid(A_n \sequoid\pr_i)\cdots))\rangle$ is given by the tree embedding 
  \[
    \phi\from P_{\prod_i (A_1 \sequoid(\cdots\sequoid(A_n\sequoid B_i)\cdots))} \to P_{A_1\sequoid(\cdots\sequoid(A_n\sequoid \prod_iB_i))}
    \]
  defined as follows.
  \[
    \phi(s) = \begin{cases}
      \epsilon & \text{if $s=\epsilon$} \\
      [\inj_{A_1,\cdots,A_n},\inj_{B_j}]_*(s) & \parbox[t][][t]{180pt}{if $s$ begins with a move in the $j$-th component}
    \end{cases}
    \]
  Note that $\phi$ is not in general injective, since if $s$ occurs entirely inside one of the copies of the $A_i$, then $\phi(s)=\phi(s')$ for any identical sequence $s'$ occurring inside one of the other copies of the $A_i$.

@@

\begin{definition}[\cite{laird02}]
  A \emph{sequoidal closed category} is an inclusive sequoidal category $\C$ such that $\C$ is a monoidal closed category (with inner hom $\implies$) and such that the map $f\mapsto \Lambda(\wk_{A,B};f)$ defines an isomorphism
  \[
    \Lambda_s \from \C_s(A \sequoid B,C) \toisom \C_s(A,B\implies C)\,.
    \]
\end{definition}

\begin{proposition}
  $\G$ is a sequoidal closed category.
\end{proposition}
\begin{proof}
  Since $\wk_{A,B}$ is an epimorphism and $\Lambda$ is a bijection, the map is certainly injective.  
  Showing that it is surjective comes down to proving that uncurrying of a strict strategy for $A\implies (B \implies C)$ is a strict strategy for $(A\sequoid B) \implies C$.
  Indeed, after the opening move in $C$, in both cases player $P$ must play the next move in $A$.
\end{proof}

\section{A Formula for the Exponential}

@@

\begin{theorem}[\cite{MelliesCofCommCom}]
  Let $\C$ be a monoidal category such that the monoidal unit for $\C$ is a terminal object.
  Suppose that $\C$ has symmetrized tensor powers that commute with the tensor product.

@@

  Then the canonical morphisms $B \tensor A^{\sequoid (n+1)}\to B\tensor A^{\sequoid n}$ are given by
  \[
    B \tensor (A \sequoid \blank)^n()\,,
    \]
  where $() \from A \to I$ is the unique morphism into the terminal object.
\end{proposition}
\begin{proof}
  First, we show by induction on $n$ that the following diagram commutes.
  \[
    \begin{tikzcd}[column sep=30pt]
      A^{\tensor(n+1)} \arrow[r, "A^{\tensor n}\tensor()"] \arrow[d, "A\tensor\wk^{n+1}"']
        & A^{\tensor n} \arrow[d, "\wk^n"] \\
      A^{\sequoid(n+1)} \arrow[r, "(A\sequoid\blank)^n()"]
        & A^{\sequoid n}
    \end{tikzcd}
    \]
  This is clearly true for $n=0$; in the general case, we have the following commutative diagram --
  \[
    \begin{tikzcd}[column sep=41pt]
      A^{\tensor(n+2)} \arrow[r] \arrow[dd, "\wk^{n+2}"']
        & A \tensor A^{\tensor (n+1)} \arrow[r, "A \tensor (A^{\tensor n}\tensor ())"] \arrow[d, "A \tensor \wk^{n+1}"']
          & A \tensor A^{\tensor n} \arrow[r] \arrow[d, "A\tensor \wk^n"]
            & A^{\tensor(n+1)} \arrow[dd, "\wk^{n+1}"] \\
      %
        & A \tensor A^{\sequoid (n+1)} \arrow[r, "A \tensor (A\sequoid\blank)^n()"] \arrow[dl, "{\wk_{A,A^{\sequoid (n+1)}}}" description]
          & A \tensor A^{\sequoid n} \arrow[dr, "{\wk_{A,A^{\sequoid n}}}" description]
            & \\
      A^{\sequoid(n+1)} \arrow[rrr, "(A \sequoid \blank)^{n+1}()"]
        &
          &
            & A^{\sequoid (n+1)}
    \end{tikzcd}\,\text{--}
    \]
  where the middle square is the inductive hypothesis (tensored by $A$), the outer trapezia are the definitions of $\wk^{n+2}$ and $\wk^{n+1}$, and the bottom trapezium commutes because $\wk$ is a natural transformation.

@@

\section{The Exponential as a Final Coalgebra}

@@

\section{Denotational Semantics of Idealized Algol}

@@

\section{The Strategy $\cell$}

@@

\section{Small-Step Operational Semantics}

@@

\section{Soundness}
